---
title: "STAT 184 Project"
author: "ProjectTeam8-4"
date: "April 20, 2021"
output: html_notebook
---

#### Load Packages
```{r}
library(tidyverse) 
library(dplyr)
```

#### Read in Data and Convert Stat Files to Data Frames
```{r}
Memorial_Fantasy <- read.csv("Memorial_Fantasy.csv") # read in Memorial_Fantasy data

names(Memorial_Fantasy) <- c("Player", "Salary", "FPTS", "Tournament", "Par") # change Name variable to Player

Memorial_Stats <- read.delim("Memorial_Stats.txt") # read in Memorial_Stats data

Memorial_Stats <- data.frame(column = c(Memorial_Stats[,]))

Memorial_Stats <- Memorial_Stats %>%
  mutate(key = rep(c("Position", "Player", "Score", "Putt", "ARG", "App", "OTT", "T2G", "Total"), n()/9), id = cumsum(key == "Position")) %>%
  spread(key, column) # long to wide transformation; transpose every 9 rows of Memorial_Stats into a column

Memorial_Stats <- subset(Memorial_Stats, select = -id)
Memorial_Stats <- Memorial_Stats[, c("Position", "Player", "Score", "Putt", "ARG", "App", "OTT", "T2G", "Total")] # reorganize columns

Schwab_Fantasy <- read.csv("Schwab_Fantasy.csv") # read in Schwab_Fantasy data

names(Schwab_Fantasy) <- c("Player", "Salary", "FPTS", "Tournament", "Par") # change Name variable to Player

Schwab_Stats <- read.delim("Schwab_Stats.txt") # read in Schwab_Stats data

Schwab_Stats <- data.frame(column = c(Schwab_Stats[,]))

Schwab_Stats <- Schwab_Stats %>%
  mutate(key = rep(c("Position", "Player", "Score", "Putt", "ARG", "App", "OTT", "T2G", "Total"), n()/9), id = cumsum(key == "Position")) %>%
  spread(key, column) # long to wide transformation; transpose every 9 rows of Memorial_Stats into a column

Schwab_Stats <- subset(Schwab_Stats, select = -id)
Schwab_Stats <- Schwab_Stats[, c("Position", "Player", "Score", "Putt", "ARG", "App", "OTT", "T2G", "Total")] # reorganize columns

PGA_Fantasy <- read.csv("PGA_Fantasy.csv") # read in PGA_Fantasy data

names(PGA_Fantasy) <- c("Player", "Salary", "FPTS", "Tournament", "Par") # change Name variable to Player

PGA_Stats <- read.delim("PGA_Stats.txt") # read in PGA_Stats data

PGA_Stats <- data.frame(column = c(PGA_Stats[,]))

PGA_Stats <- PGA_Stats %>%
  mutate(key = rep(c("Position", "Player", "Score", "Putt", "ARG", "App", "OTT", "T2G", "Total"), n()/9), id = cumsum(key == "Position")) %>% 
  spread(key, column) # long to wide transformation; transpose every 9 rows of Memorial_Stats into a column

PGA_Stats <- subset(PGA_Stats, select = -id)
PGA_Stats <- PGA_Stats[, c("Position", "Player", "Score", "Putt", "ARG", "App", "OTT", "T2G", "Total")] # reorganize columns
```

#### Join Fantasy and Statistics to Create One Data Frame for Each of Three Tournaments
```{r}
Memorial <- full_join(Memorial_Stats, Memorial_Fantasy, by = "Player")

PGA <- full_join(PGA_Stats, PGA_Fantasy, by = "Player")
  
Schwab <- full_join(Schwab_Stats, Schwab_Fantasy, by = "Player")
```

#### Combine the Three Tournament Data Frames to Final Data Frame

This last R chunk combines the 3 data frames above--Memorial, PGA and Schwab--while removing the 4 rows that are mainly filled with NA values and replacing the score E with 0. Next, we created the additional 9 variables by mutating the variables we already had. We created the top10 and MadeCut variables by using an if else statement to see if each grepl condition was met. To create the SalaryRange variable we used the function case_when. This allowed us to pick specific conditions and, if they were met, they were mutated to the desired range. For the rank variable we used an if else statement with a grepl condition. Depending if the statement was true or false, it went through a gsub where the values were changed to a number or NA. For the avg variable we made sure the Charles Schwab, PGA Championship and Memorial variables were numeric. By doing this, we were able to take the sum and divide it by the tournplayed variable to receive the average. For the tournplayed variable we used gsub for each of the 3 tournaments to replace it with a value of 1. Again, we made sure it was numeric so that we could add them up to get the total tournplayed. If an NA was present in the Charles Schwab, PGA Championship or Memorial variable, that player would not be credited for participating in that respective tournament. Lastly, throughout the code we rearranged the variables so that it would be in the same order as presented to us in the project description.

```{r}
Mem_PGA_Sch <- rbind(Memorial, PGA, Schwab) # stack the 3 data frames together

Mem_PGA_Sch <- Mem_PGA_Sch[complete.cases(Mem_PGA_Sch[1:9]),] %>% # removes 4 rows with NA
  mutate(Score = ifelse(as.character(Score) == "E", 0, as.character(Score))) # replace score E with 0

Mem_PGA_Sch <- Mem_PGA_Sch %>%
  mutate(top10= if_else(grepl("^[1-9]$|10|^T[1-9]$|T10", Position),"Yes", "No")) # add top10 variable

Mem_PGA_Sch <- Mem_PGA_Sch %>%
  mutate(MadeCut = if_else(grepl("CUT", Position),"No","Yes")) # add MadeCut variable

Temp_Mem_PGA_Sch <- Mem_PGA_Sch %>%
  group_by(Player) %>%
  select(Player, Score, Tournament)

Temp_Mem_PGA_Sch <- pivot_wider(Temp_Mem_PGA_Sch, names_from = Tournament, values_from = Score)

Temp_Mem_PGA_Sch <- Temp_Mem_PGA_Sch[, c("Player", "Charles Schwab", "Memorial", "PGA Championship")] # reorganize columns

Mem_PGA_Sch <- Mem_PGA_Sch %>%
  left_join(Temp_Mem_PGA_Sch, by = "Player")

Mem_PGA_Sch <- Mem_PGA_Sch %>%
  mutate(SalaryRange = case_when(grepl("6[0-9]{3}", Salary) ~ "6000s", grepl("7[0-9]{3}", Salary) ~ "7000s", grepl("8[0-9]{3}", Salary) ~ "8000s",grepl("9[0-9]{3}", Salary) ~ "9000s",grepl("1[0-9]{4}", Salary) ~ "10000s")) # add SalaryRange variable

Mem_PGA_Sch <- Mem_PGA_Sch[, c("Position", "Player", "Score", "Putt", "ARG", "App", "OTT", "T2G", "Total", "Salary", "FPTS", "Tournament", "Par", "top10", "MadeCut", "SalaryRange", "Charles Schwab", "Memorial", "PGA Championship")] # reorganize columns

Temp_Mem_PGA_Sch_2 <- Mem_PGA_Sch %>%
  select(Position, Player, Tournament, `Charles Schwab`, Memorial, `PGA Championship`) %>%
  mutate(playedSchwab = gsub(pattern = "^-\\d{2}|^\\d{2}|^-\\d{1}|^\\d{1}|", replacement = "1", `Charles Schwab`)) %>%
  mutate(playedMemorial = gsub(pattern = "^-\\d{2}|^\\d{2}|^-\\d{1}|^\\d{1}|", replacement = "1", Memorial)) %>%
  mutate(playedPGA = gsub(pattern = "^-\\d{2}|^\\d{2}|^-\\d{1}|^\\d{1}|", replacement = "1", `PGA Championship`)) %>%
  mutate(playedSchwab = as.numeric(playedSchwab)) %>%
  mutate(playedMemorial = as.numeric(playedMemorial)) %>%
  mutate(playedPGA = as.numeric(playedPGA)) %>%
  rowwise() %>%
  mutate(tournplayed = sum(playedSchwab, playedMemorial, playedPGA, na.rm = TRUE)) # create tournplayed variable

Temp_Mem_PGA_Sch_2 <- subset(Temp_Mem_PGA_Sch_2, select = -c(`Charles Schwab`, Memorial, `PGA Championship`, playedSchwab, playedMemorial, playedPGA)) # delete extra columns

Mem_PGA_Sch <- Mem_PGA_Sch %>%
  left_join(Temp_Mem_PGA_Sch_2, by = c("Position", "Player", "Tournament")) # add tournplayed variable

Mem_PGA_Sch <- Mem_PGA_Sch %>%
  relocate(tournplayed, .after = SalaryRange) # reorganize columns

Mem_PGA_Sch <- Mem_PGA_Sch %>%
  mutate(`Charles Schwab` = as.numeric(`Charles Schwab`)) %>%
  mutate(Memorial = as.numeric(Memorial)) %>%
  mutate(`PGA Championship` = as.numeric(`PGA Championship`)) %>%
  rowwise() %>%
  mutate(avg = sum(`Charles Schwab`, Memorial, `PGA Championship`, na.rm = TRUE)/tournplayed) # add avg variable
  
Mem_PGA_Sch$avg <- round(Mem_PGA_Sch$avg, digit = 1) # round avg to 1 decimal place

Mem_PGA_Sch <-
  Mem_PGA_Sch %>%
  mutate(rank = if_else(grepl("^T",Position), gsub("^T", "",Position), gsub("CUT", "NA", Position))) # add rank variable

FinalDataFrame <- Mem_PGA_Sch %>%
  relocate(rank, .after = Par) # organize columns
```

#### Research Question One
```{r}
#Possible Research Questions
  #1  What is the distribution of salaries for the top 20 golfers?
  #2  What is the distribution of scores for the top 10 ranked golfers?
  #3  What is the correlation between the T2G and the rank of a player in the Charles Schwab Cup?
  #4  Is there a correlation between a player’s rank and whether or not they made the cut?
  #5  What is the distribution of players who Made the Cut with only 1 tournament played for all other players who Made the Cut?
  #6  What are the top 10 golfers average scores for the three different tournaments?
  #7  What was the average score for the three tournaments of golfers in ranks 40-50?
  #8  Which tournament has the most number of golfers entered?
  #9  What distribution of golfers posted a final score of ±0 in the Memorial Tournament?
  #10 What are the salaries of the top 10 ranked golfers?  
```

#### Reaserch Question Two
```{r}

```

