---
title: "STAT 184 Project"
author: "ProjectTeam8-4"
date: "May 3, 2021"
output: html_notebook 
---

#### Load Packages
```{r message=FALSE}
library(tidyverse)
```

#### Read in Data and Convert Stat files to data frames
```{r}
Memfan <- read.csv("Memorial_Fantasy.csv", header = T)
Pgafan <- read.csv("PGA_Fantasy.csv", header = T)
Schfan <- read.csv("Schwab_Fantasy.csv", header = T)
Memstat<- read.delim("Memorial_Stats.txt", header = T )
Memstats <-  as.data.frame(matrix(Memstat$Data, ncol=9, byrow=TRUE))
Pgastat<- read.delim("PGA_Stats.txt", header = T)
Pgastats <-  as.data.frame(matrix(Pgastat$Data, ncol=9, byrow=TRUE))
Schstat<- read.delim("Schwab_Stats.txt", header = T)
Schstats <-  as.data.frame(matrix(Schstat$Data, ncol=9, byrow=TRUE))
```

#### Join Fantasy and Statistics to Create One Data Frame for Each of Three Tournaments

```{r}
Memorial <- Memstats %>%
  inner_join(Memfan, by=c("V2"= "Name"))
Pga <- Pgastats %>%
  inner_join(Pgafan, by=c("V2"= "Name")) 
Schwab <- Schstats %>%
  inner_join(Schfan, by=c("V2"= "Name"))
```

#### Combine the Three Tournament Data Frames to Final Data Frame
Include paragraph explaining what the R statements are "doing" and what you are trying to accomplish with these statements.

```{r}
All <- rbind(Memorial, Pga, Schwab)
names(All)[1:9] <- c('Position', 'Player', 'Score', 
                          'Putt', 'ARG', 'App', 'OTT', 'T2G', 'Total')
All$Score <- fct_recode(All$Score, "0" = "E") 
Top10 <- 
  All %>%
  group_by(Tournament) %>%
  mutate(Score = as.numeric(as.character(factor(Score)))) %>%
  mutate(rank = rank(Score, ties.method = "min")) %>%
  mutate(top10 = if_else(rank <= 10, 'Yes', 'No')) %>%
  mutate(MadeCut = if_else(Position =='CUT', 'No', 'Yes')) %>%
  mutate(SalaryRange = ifelse(grepl("^6", Salary), "6000s",
                       ifelse(grepl("^7", Salary), "7000s",
                       ifelse(grepl("^8", Salary), "8000s",
                       ifelse(grepl("^9" , Salary), "9000s",
                       ifelse(grepl("^10", Salary), "10000s", "11000s"))))))
TournPlayed <- All %>%
  select((Player)) %>%
  group_by(Player) %>%
  mutate(tournplayed = n()) %>%
  distinct()
AvgScore <- All %>%
    select(Player, Score, Tournament) %>%
    mutate(Score = as.numeric(as.character(factor(Score)))) %>%
    spread(Tournament, Score) %>%
    mutate(avg = round(rowMeans(.[2:4], na.rm=TRUE), 1))
Results_Player <- Top10 %>%
  left_join(TournPlayed, by = "Player")
AllResults <- Results_Player %>%
  left_join(AvgScore, by = "Player")
```

#### Research Question One

Depending on whether a player made the cut or not, is there a correlation between a player's rank in a tournament and the amount of fantasy points they generate in that same tournament?

```{r}
RQ1 <-
  AllResults%>%
  mutate(MadeCut = recode(MadeCut, "Yes" = "Made Cut: Yes", "No"= "Made Cut: No"))

ggplot(RQ1)+
  aes(x = rank, y = FPTS, colour = Tournament)+
  geom_point(size=1L)+
  xlab("Rank in Tournament")+
  ylab("Fantasy Points (FPTS)")+
  scale_x_reverse() +
  scale_color_hue() +
  theme_bw() +
  facet_wrap(vars(MadeCut))
```
To answer this question, we used a scatter plot to determine if there was a correlation between a player's rank and the amount of fantasy points they generated. The best players are ranked higher. For example, the player ranked number one is the highest ranked player. As depicted in the graph, among players who made the cut, there is a very strong, positive correlation between a player's rank and the amount of fantasy points they acquired. This positive correlation between rank and fantasy points is also reflected among players who did not make the cut. This suggests that, regardless of whether a player made the cut or not, a higher rank correlates to more fantasy points. The best ranked players, such as those in the Top 10, earn significantly more fantasy points than those players in the middle of the pack. These graphs also suggest that player's typically gained more fantasy points in the Charles Schwab and PGA Championship tournaments than in the Memorial tournament. All in all, a higher rank correlates strongly to generating more fantasy points.

#### Research Question Two

What is the relationship in each tournament between the scores of the players who were in the Top 10 and the scores of the players not in the Top 10?

```{r}
ggplot(AllResults)+
  aes(x = top10, y = Score, fill = Tournament) +
  geom_boxplot() +
  xlab("Top 10 in Tournament") +
  ylab("Scores in Tournament") +
  scale_fill_hue() +
  theme_bw() +
  facet_wrap(vars(Tournament))
```
To answer this question we chose to use a box plot. As expected, the scores of players who were in the Top 10 were much lower (better) than the scores of players who were not in the Top 10 for all three tournaments--Charles Schwab, Memorial and PGA Championship. The median scores of players who were not in the Top 10 were roughly between 0 and 6. Players not within the Top 10 performed the best in terms of score within the Charles Schwab tournament, followed by the PGA Championship and then the Memorial. All three tournaments had an at least one outlier in the positive direction among players who did not make the Top 10, showing players who performed especially poor in terms of score within that tournament. Amongst players in the Top 10, golfers once again achieved their best scores in the Charles Schwab tournament, followed by the PGA Championship and then the Memorial. While the Memorial tournament held the worst/highest scores for both players in and out of the Top 10, this tournament also had the smallest spread and deviation between these two groups. The only outlier among players in the Top 10 occurred in the PGA Championship, representing a player who performed incredibly well relative to his competition at the tournament. Overall, the scores of players who were in Top 10 were higher than the scores of players who were not in the Top 10 for all three tournaments, and players typically performed the best in terms of score at the Charles Schwab Tournament which had the lowest/best median score for players in both the Top 10 and not in the Top 10 (as well as the lowest/best score overall and the lowest/best maximum score).


 