---
title: "Stat 184 Project"
author: "ProjectTeam8-4"
date: "May 3, 2021"
output: html_notebook 
---

#### Load Packages
```{r message=FALSE}
library(tidyverse) 
```

#### Read in Data and Convert Stat files to data frames
```{r}
Memfan <- read.csv("Memorial_Fantasy.csv", header = T)
Pgafan <- read.csv("PGA_Fantasy.csv", header = T)
Schfan <- read.csv("Schwab_Fantasy.csv", header = T)

Memstat<- read.delim("Memorial_Stats.txt", header = T )
Memstats <-  as.data.frame(matrix(Memstat$Data, ncol=9, byrow=TRUE))

Pgastat<- read.delim("PGA_Stats.txt", header = T)
Pgastats <-  as.data.frame(matrix(Pgastat$Data, ncol=9, byrow=TRUE))

Schstat<- read.delim("Schwab_Stats.txt", header = T)
Schstats <-  as.data.frame(matrix(Schstat$Data, ncol=9, byrow=TRUE))
```

#### Join Fantasy and Statistics to Create One Data Frame for Each of Three Tournaments

```{r}
Memorial <- Memstats %>%
  inner_join(Memfan, by=c("V2"= "Name"))

Pga <- Pgastats %>%
  inner_join(Pgafan, by=c("V2"= "Name")) 


Schwab <- Schstats %>%
  inner_join(Schfan, by=c("V2"= "Name"))
```

#### Combine the Three Tournament Data Frames to Final Data Frame
Include paragraph explaining what the R statements are "doing" and what you are trying to accomplish with these statements.

```{r}
All <- rbind(Memorial, Pga, Schwab)
names(All)[1:9] <- c('Position', 'Player', 'Score', 
                          'Putt', 'ARG', 'App', 'OTT', 'T2G', 'Total')

All$Score <- fct_recode(All$Score, "0" = "E") 

Top10 <- 
  All %>%
  group_by(Tournament) %>%
  mutate(Score = as.numeric(as.character(factor(Score)))) %>%
  mutate(rank = rank(Score, ties.method = "min")) %>%
  mutate(top10 = if_else(rank <= 10, 'Yes', 'No')) %>%
  mutate(MadeCut = if_else(Position =='CUT', 'No', 'Yes')) %>%
  mutate(SalaryRange = ifelse(grepl("^6", Salary), "6000s",
                       ifelse(grepl("^7", Salary), "7000s",
                       ifelse(grepl("^8", Salary), "8000s",
                       ifelse(grepl("^9" , Salary), "9000s",
                       ifelse(grepl("^10", Salary), "10000s", "11000s"))))))

TournPlayed <- All %>%
  select((Player)) %>%
  group_by(Player) %>%
  mutate(tournplayed = n()) %>%
  distinct()

AvgScore <- All %>%
    select(Player, Score, Tournament) %>%
    mutate(Score = as.numeric(as.character(factor(Score)))) %>%
    spread(Tournament, Score) %>%
    mutate(avg = round(rowMeans(.[2:4], na.rm=TRUE), 1))

Results_Player <- Top10 %>%
  left_join(TournPlayed, by = "Player")

AllResults <- Results_Player %>%
  left_join(AvgScore, by = "Player")
```

#### Research Question One
```{r}

```

#### Research Question Two
```{r}

```

